---
title: "AQI shiny dashboard README"
author: "Spencer Zhang"
date: "3/11/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, eval = F)
```

```{r library, eval = T}
library(jsonlite)
library(tidyverse)
library(sf)
library(tigris)
library(censusapi)
library(mapview)
library(leaflet)
library(mapboxapi)
library(readr)
```

# Geography analysis

The air quality data, specifically PM2.5 daily concentration, is from sensor data by [PurpleAir](https://www2.purpleair.com). The sensors are labeled as "outside" and "inside", indicating applicaiton of sensors for outdoor air quality and indoor air quality.
In this analysis, we focus on San Mateo County. The data for Redwood City and San Mateo City is processed by the author. The data for other cities are kindly supported to this dashboard by my colleagues in Stanford CEE218Y, W22. The data for East Palo Alto and Foster City is processed by Daphne, Lena, and Catherine, and the Belmont data is processed by Awoe.
```{r pa-api-reading}
pa_api <- "CAD01D42-A086-11EC-B9BF-42010A800003"

json <- fromJSON(paste0(
    "https://api.purpleair.com/v1/sensors?api_key=",
    pa_api,
    "&fields=name,location_type,latitude,longitude,pm2.5_1week,temperature,humidity,primary_id_a,primary_key_a,secondary_id_a,secondary_key_a,primary_id_b,primary_key_b,secondary_id_b,secondary_key_b"
  ))

all_sensors <- json %>% 
  .$data %>% 
  as.data.frame() %>% 
  set_names(json$fields) %>% 
  filter(
    !is.na(longitude),
    !is.na(latitude)
  ) %>% 
  st_as_sf(coords = c("longitude","latitude"), crs = 4326) %>% 
  mutate(location_type = ifelse(
    location_type == 0,
    "outside",
    "inside"
  ))

#test
view(head(all_sensors))
```

```{r test-read-fc-data-from-lena}
#read FC daily average PM2.5 data in Feb 2022
fc <- read_rds("fc_sensor_data.rds")

plot(fc$PM25)
```

```{r smc-boundary}
smc_boundary <- counties(state = "CA", cb= T) %>%
  filter(NAME == "San Mateo")
```

```{r sensors-in-smc}
smc_sensors <-
  all_sensors %>% 
  .[smc_boundary %>% st_transform(4326), ] #1079 sensors in SMC 
```

## AQI data processing for Redwood City and San Mateo City

In total, Redwood City has 105 sensors and San Mateo City has 108 sensors. The sensors are disapportionaly distributed according to the choice of the purchased customers. In order to make the data representative for the City, we used voronoi technique to assign the area of the city to its closet sensors. The final PM2.5 concentration is a weighted average of the sensor data by its assigned adjacent area. The voronoi technique is applied seperately for indoor and outdoor sensors.
```{r rwc&smc-sensors}
rwc_boundary <- places(state = "CA", cb = T) %>%
  filter(NAME == "Redwood City")

rwc_sensors <-
  all_sensors %>% 
  .[rwc_boundary%>% st_transform(4326), ] 

sm_boundary <- places(state = "CA", cb = T) %>%
  filter(NAME == "San Mateo")

sm_sensors <-
  all_sensors %>% 
  .[sm_boundary%>% st_transform(4326), ]

#test view
leaflet() %>%
  addTiles() %>%
  addPolygons(
    data = rwc_boundary
  )
```

```{r tester}
start <- "2022-02-01%2000:08:00"
end <- "2022-02-10%2000:08:00"

row <- 1
print(paste0(row,". ",rwc_sensors[row,]$sensor_index))

  a1 <- read_csv(paste0(
    "https://api.thingspeak.com/channels/",
    rwc_sensors[row,]$primary_id_a,
    "/feeds.csv?api_key=",
    rwc_sensors[row,]$primary_key_a,
    "&average=1440&round=3&start=",start,
    "&end=", end, 
    "&timezone=America/Los_Angeles"
  ), show_col_types = F) %>% 
    set_names(c("created_at","PM1.0_CF_1_ug/m3_A","PM2.5_CF_1_ug/m3_A","PM10.0_CF_1_ug/m3_A","Uptime_Minutes_A","RSSI_dbm_A","Temperature_F_A","Humidity_%_A","PM2.5_CF_ATM_ug/m3_A"))
  
    a1_ave <- read_csv(paste0(
    "https://api.thingspeak.com/channels/",
    rwc_sensors[row,]$primary_id_a,
    "/feeds.csv?api_key=",
    rwc_sensors[row,]$primary_key_a,
    "&average=1440&round=3&start=",start,
    "&end=", end, 
    "&timezone=America/Los_Angeles&average=daily"
  ), show_col_types = F) %>% 
    set_names(c("created_at","PM1.0_CF_1_ug/m3_A","PM2.5_CF_1_ug/m3_A","PM10.0_CF_1_ug/m3_A","Uptime_Minutes_A","RSSI_dbm_A","Temperature_F_A","Humidity_%_A","PM2.5_CF_ATM_ug/m3_A"))
```

```{r rwc-sensor-voronoi}
rwc_pm25_out_voronoi <-
  rwc_sensors %>%
  filter(location_type == "outside") %>% 
  st_union() %>% 
  st_voronoi() %>% 
  st_cast() %>% 
  st_as_sf() %>% 
  st_intersection(.,st_union(rwc_boundary %>% st_transform(4326))) %>% 
  st_join(rwc_sensors %>% filter(location_type == "outside"))

rwc_pm25_in_voronoi <-
  rwc_sensors %>%
  filter(location_type == "inside") %>% 
  st_union() %>% 
  st_voronoi() %>% 
  st_cast() %>% 
  st_as_sf() %>% 
  st_intersection(.,st_union(rwc_boundary %>% st_transform(4326))) %>% 
  st_join(rwc_sensors %>% filter(location_type == "inside"))

ggplot(rwc_pm25_in_voronoi) + geom_sf()  
```

```{r sm-sensor-voronoi}
sm_pm25_out_voronoi <-
  sm_sensors %>%
  filter(location_type == "outside") %>% 
  st_union() %>% 
  st_voronoi() %>% 
  st_cast() %>% 
  st_as_sf() %>% 
  st_intersection(.,st_union(sm_boundary %>% st_transform(4326))) %>% 
  st_join(sm_sensors %>% filter(location_type == "outside"))

sm_pm25_in_voronoi <-
  sm_sensors %>%
  filter(location_type == "inside") %>% 
  st_union() %>% 
  st_voronoi() %>% 
  st_cast() %>% 
  st_as_sf() %>% 
  st_intersection(.,st_union(sm_boundary %>% st_transform(4326))) %>% 
  st_join(sm_sensors %>% filter(location_type == "inside"))

ggplot(sm_pm25_out_voronoi) + geom_sf()
ggplot(sm_pm25_in_voronoi) + geom_sf()
```

```{r sensor-weighted}
#County block groups
smc_cbgs <- block_groups("CA","San Mateo County", cb = T, progress_bar = F) %>% 
  st_transform(4326)

rwc_pm25_voronoi_cbg <-
  rwc_pm25_voronoi %>% 
  st_intersection(smc_cbgs) %>% 
  st_make_valid() %>% 
  mutate(
    area = st_area(.) %>% as.numeric()
  ) %>% 
  st_drop_geometry() %>% 
  group_by(GEOID) %>% 
  summarize(
    PM25 = weighted.mean(PM25, area, na.rm = T)
  ) %>% 
  left_join(rwc_cbgs %>% dplyr::select(GEOID)) %>% 
  st_as_sf()

rwc_sensors_out_weighted <-
  rwc_pm25_out_voronoi %>% 
  mutate(
    area = st_area(.) %>% as.numeric()
  ) 

rwc_sensors_in_weighted <-
  rwc_pm25_in_voronoi %>% 
  mutate(
    area = st_area(.) %>% as.numeric()
  ) 

sm_sensors_out_weighted <-
  sm_pm25_out_voronoi %>% 
  mutate(
    area = st_area(.) %>% as.numeric()
  ) 

sm_sensors_in_weighted <-
  sm_pm25_in_voronoi %>% 
  mutate(
    area = st_area(.) %>% as.numeric()
  ) 

sm_cbgs <- block_groups("CA","San Mateo City", cb = T, progress_bar = F) %>% 
  st_transform(4326)

sm_pm25_voronoi_cbg <-
  sm_pm25_voronoi %>% 
  st_intersection(sm_cbgs) %>% 
  st_make_valid() %>% 
  mutate(
    area = st_area(.) %>% as.numeric()
  ) %>% 
  st_drop_geometry() %>% 
  group_by(GEOID) %>% 
  summarize(
    PM25 = weighted.mean(PM25, area, na.rm = T)
  ) %>% 
  left_join(sm_cbgs %>% dplyr::select(GEOID)) %>% 
  st_as_sf()
```

We processed the daily average PM2.5 concentration for the time period of February 2022 in all cities.
```{r rwc-data-download}
time_stamp <- c(
  "2022-02-01%2000:08:00",
  "2022-02-11%2000:08:00",
  "2022-02-21%2000:08:00",
  "2022-02-28%2000:08:00"
)

#change this and run temp code repeatedly without modifying the loop code
sensors_weighted <- rwc_sensors_in_weighted

temp <- 
  1:3 %>%
  map_dfr(function(x){
    #loop through time periods
    start <- time_stamp[x]
    end <- time_stamp[x+1]
    print(paste0("now running time from ", start, " to ", end))
    
    data <-
      1:nrow(sensors_weighted) %>% 
      map_dfr(function(row){
      
      print(paste0(row,". ",sensors_weighted[row,]$sensor_index))
      
      a1 <- read_csv(paste0(
        "https://api.thingspeak.com/channels/",
        sensors_weighted[row,]$primary_id_a,
        "/feeds.csv?api_key=",
        sensors_weighted[row,]$primary_key_a,
        "&average=1440&round=3&start=",start,
        "&end=", end, 
        "&timezone=America/Los_Angeles"
      ), show_col_types = F) %>% 
        set_names(c("created_at","PM1.0_CF_1_ug/m3_A","PM2.5_CF_1_ug/m3_A","PM10.0_CF_1_ug/m3_A","Uptime_Minutes_A","RSSI_dbm_A","Temperature_F_A","Humidity_%_A","PM2.5_CF_ATM_ug/m3_A"))
          
      a2 <- read_csv(paste0(
        "https://api.thingspeak.com/channels/",
        sensors_weighted[row,]$secondary_id_a,
        "/feeds.csv?api_key=",
        sensors_weighted[row,]$secondary_key_a,
        "&average=1440&round=3&start=",start,
        "&end=", end, 
        "&timezone=America/Los_Angeles"
      ), show_col_types = F) %>% 
        set_names(c("created_at","0.3um/dl_A","0.5um/dl_A","1.0um/dl_A","2.5um/dl_A","5.0um/dl_A","10.0um/dl_A","PM1.0_CF_ATM_ug/m3_A","PM10_CF_ATM_ug/m3_A"))
        
      b1 <- read_csv(paste0(
        "https://api.thingspeak.com/channels/",
        sensors_weighted[row,]$primary_id_b,
        "/feeds.csv?api_key=",
        sensors_weighted[row,]$primary_key_b,
        "&average=1440&round=3&start=",start,
        "&end=", end, 
        "&timezone=America/Los_Angeles"
      ), show_col_types = F) %>% 
        set_names(c("created_at","PM1.0_CF_1_ug/m3_B","PM2.5_CF_1_ug/m3_B","PM10.0_CF_1_ug/m3_B","HEAP_B","ADC0_voltage_B","Atmos_Pres_B","Not_Used_B","PM2.5_CF_ATM_ug/m3_B"))
      
      b2 <- read_csv(paste0(
        "https://api.thingspeak.com/channels/",
        sensors_weighted[row,]$secondary_id_b,
        "/feeds.csv?api_key=",
        sensors_weighted[row,]$secondary_key_b,
        "&average=1440&round=3&start=",start,
        "&end=", end, 
        "&timezone=America/Los_Angeles"
      ), show_col_types = F) %>% 
        set_names(c("created_at","0.3um/dl_B","0.5um/dl_B","1.0um/dl_B","2.5um/dl_B","5.0um/dl_B","10.0um/dl_B","PM1.0_CF_ATM_ug/m3_B","PM10_CF_ATM_ug/m3_B"))
      
      combined <- a1 %>% 
        left_join(a2, by = "created_at") %>% 
        left_join(b1, by = "created_at") %>% 
        left_join(b2, by = "created_at") %>% 
        transmute(
          date = as.Date(created_at),
          ID = as.numeric(sensors_weighted[row,]$sensor_index),
          Location = sensors_weighted[row,]$location_type,
          PM25 = 0.524*as.numeric(`PM2.5_CF_1_ug/m3_A`) - 0.0852*as.numeric(`Humidity_%_A`) + 5.72,
          PM25_CF1_A = as.numeric(`PM2.5_CF_1_ug/m3_A`),
          PM25_CF1_B = as.numeric(`PM2.5_CF_1_ug/m3_B`),
          PM25_ATM_A = as.numeric(`PM2.5_CF_ATM_ug/m3_A`),
          PM25_ATM_B = as.numeric(`PM2.5_CF_ATM_ug/m3_B`),
          Humidity = as.numeric(`Humidity_%_A`)
        ) 
    })
    
    saveRDS(data, paste0("rwc_indoors_raw", x, ".rds"))
    return(data)
  }
  )

view(head(temp))

saveRDS(temp, "rwc_indoor_raw.rds")

```

```{r read-raw-data}

sm_outdoor_raw <- read_rds("sm_outdoor_raw.rds")
sm_indoor_raw <- read_rds("sm_indoor_raw.rds")
rwc_indoor_raw <- read_rds("rwc_indoor_raw.rds")
rwc_outdoor_raw <- read_rds("rwc_outdoor_raw.rds")
```

According to [PurpleAir's documentation](https://docs.google.com/document/d/15ijz94dXJ-YAZLi9iZ_RaBwrZ4KtYeCy08goGBwnbCU/edit#), the data column labelled "CF=1" is calibrated for indoor application and "CF=ATM" is calibrated for outdoor application. Therefore, we used the former one for inside AQI/PM2.5 calculation and latter for outside, shown in the final dashboard map.
 
For data quality control, there are two channels (A and B) in each sensor to maintain quality control. If both laser counters (channels) are in agreement, the data can be seen as excellent quality. If there are different readings from the two channels, there may be a fault with one or both. Therefore, we conduct quality control by inspecting the difference between channels for sensors in Redwood City and San Mateo City.
There are two outside sensors in Redwood City with severe quality issues, ID = 
The inside sensors in Redwood City are all of excellent data quality.

For San Mateo City, most of the inside sensors only has data for Channel A, so we skip the difference comparison step and look at the distribution of observed data in Channel A alone. There are no anomaly in the observed data.
For outside sensors in San Mateo City, the same applies for the missing data on Channel B. However the observed data in Channel A.
 
```{r data-cleansing}
#use this data frame as temp to run following code
#repeat with rwc_in, rwc_out, sm_in, sm_out
raw_data <- sm_outdoor_raw

names(raw_data)
plot(raw_data$PM25_CF1_A)
plot(raw_data$PM25_ATM_A)
plot(raw_data$PM25_CF1_B) #>4500
plot(raw_data$PM25_ATM_B) #>3000


test <- raw_data %>%
  mutate(
    cf1dif = PM25_CF1_A - PM25_CF1_B,
    amtdif = PM25_ATM_A - PM25_ATM_B,
    PM_CF1 = 0.524*(PM25_CF1_A + PM25_CF1_B)/2 - 0.0852*Humidity + 5.72,
    PM_ATM =  0.524*(PM25_ATM_A + PM25_ATM_B)/2 - 0.0852*Humidity + 5.72
  )

sum(is.na(test$PM_CF1))
sum(!is.na(test$PM_CF1))

plot(test$cf1dif) 
#RWC OUT: 6 cluster of data difference is way high (>4000)
#RWC IN: no big different
#SM IN: one cluster < -15
plot(test$amtdif) 
#RWC OUT: 6 cluster of data is way high (>3000)
#RWC IN: no big different
names(test)

sm_out_fault <- raw_data %>%
  filter(PM25_CF1_B > 4500 | PM25_ATM_B > 3000) %>%
  group_by(ID) %>%
  summarise(count=n(),
            PM25_CF1_dif = mean(PM25_CF1_A - PM25_CF1_B),
            PM25_ATM_dif = mean(PM25_ATM_A - PM25_ATM_B)
    ) %>%
  mutate(ID = as.character(ID)) %>%
  left_join(sm_sensors %>% select(
    ID = sensor_index, name
  ))

rwc_out_fault <- test %>%
  filter(cf1dif > 4000 | cf1dif < -4000 | amtdif > 3000 | amtdif < -3000) %>%
  group_by(ID) %>%
  summarise(count = n(), PM25_CF1_dif = mean(cf1dif), PM25_ATM_dif = mean(amtdif)) %>%
  mutate(ID = as.character(ID)) %>%
  left_join(rwc_sensors %>% select(
    ID = sensor_index, name
  ))

fault_sensors <- rwc_out_fault %>% mutate(
  City = "Redwood City",
  Location = "outside"
) %>%
  rbind(sm_out_fault %>% mutate(
  City = "San Mateo City",
  Location = "outside"
))

saveRDS(fault_sensors, "fault_sensors.rds")
```

The four faulty sensors excluded from the following analysis is shown in Map 1.
```{r faulty-sensor-info, eval = T}
fault_sensors <- read_rds("fault_sensors.rds")
knitr::kable(fault_sensors,
             caption = "The information of faulty sensors in RWC and SMC (count indicates the number of observations shown high variances between Channel A and B)")
```

```{r faulty-sensor-map, eval = T, fig.cap="Figure 1. The locations of faulty sensors in Redwood City and San Mateo City."}
mapview(fault_sensors %>% st_as_sf(), zcol = "name", layer.name = "Faulty sensor")
```

The voronoi technique, i.e. area weight for each sensor, is reapplied after excluding faulty sensors.
```{r adjust-voronoi-by-excluding-faulty-sensors}
#remove 2 outside sensors in rwc
fault_sensors$ID

rwc_sensors_out_weighted_corrected <-
  rwc_sensors %>%
  filter(location_type == "outside" & !sensor_index %in% fault_sensors$ID) %>% 
  st_union() %>% 
  st_voronoi() %>% 
  st_cast() %>% 
  st_as_sf() %>% 
  st_intersection(.,st_union(rwc_boundary %>% st_transform(4326))) %>% 
  st_join(rwc_sensors %>% filter(location_type == "outside"& !sensor_index %in% fault_sensors$ID)) %>% 
  mutate(
    area = st_area(.) %>% as.numeric()
  )

#remove 2 outside sensors in smc
sm_sensors_out_weighted_corrected <-
  sm_sensors %>%
  filter(location_type == "outside" & !sensor_index %in% fault_sensors$ID) %>% 
  st_union() %>% 
  st_voronoi() %>% 
  st_cast() %>% 
  st_as_sf() %>% 
  st_intersection(.,st_union(sm_boundary %>% st_transform(4326))) %>% 
  st_join(sm_sensors %>% filter(location_type == "outside"& !sensor_index %in% fault_sensors$ID)) %>% 
  mutate(
    area = st_area(.) %>% as.numeric()
  )
```

```{r calculate-weighted-daily-average}
colSums(is.na(rwc_outdoor_weighted))

rwc_outdoor_weighted <- rwc_sensors_out_weighted_corrected %>%
      st_drop_geometry() %>%
      select(
        sensor_index, area
      ) %>%
  left_join(
    rwc_outdoor_raw %>%
      mutate(sensor_index = as.character(ID))
  ) %>%
  group_by(date, ID) %>%
  summarize_all(first) %>% #remove duplicate observations of the same day by looping start&end defination 
  mutate(
    PM25 = 0.524*(PM25_ATM_A + PM25_ATM_B)/2 - 0.0852*Humidity + 5.72 #use CF = ATM for outside data
  ) %>%
  group_by(date, Location) %>%
  summarize(PM25 = weighted.mean(PM25, area, na.rm = T)) %>%
  mutate(city = "Redwood City")

sm_outdoor_weighted <- sm_sensors_out_weighted_corrected %>%
      st_drop_geometry() %>%
      select(
        sensor_index, area
      ) %>%
  left_join(
    sm_outdoor_raw %>%
      mutate(sensor_index = as.character(ID))
  ) %>%
  group_by(date, ID) %>%
  summarize_all(first) %>% #remove duplicate observations of the same day by looping start&end defination 
  mutate(
    PM25 = 0.524*(PM25_ATM_A + PM25_ATM_B)/2 - 0.0852*Humidity + 5.72 #use CF = ATM for outside data
  ) %>%
  group_by(date, Location) %>%
  summarize(PM25 = weighted.mean(PM25, area, na.rm = T)) %>%
  mutate(city = "San Mateo City")

rwc_indoor_weighted <- rwc_indoor_raw %>%
  mutate(sensor_index = as.character(ID)) %>%
  left_join(
    rwc_sensors_in_weighted %>%
      st_drop_geometry() %>%
      select(
        sensor_index, area
      )) %>%
  group_by(date, ID) %>%
  summarize_all(first) %>%
  group_by(date, Location) %>%
  summarize(PM25 = weighted.mean(PM25, area, na.rm = T)) %>%
  mutate(city = "Redwood City")

sm_indoor_weighted <- sm_indoor_raw %>%
  mutate(sensor_index = as.character(ID)) %>%
  left_join(
    sm_sensors_in_weighted %>%
      st_drop_geometry() %>%
      select(
        sensor_index, area
      )) %>%
    group_by(date, ID) %>%
  summarize_all(first) %>%
  group_by(date, Location) %>%
  summarize(PM25 = weighted.mean(PM25, area, na.rm = T)) %>%
  mutate(city = "San Mateo City")

saveRDS(rwc_outdoor_weighted, "rwc_outdoor_weighted.rds")
saveRDS(rwc_indoor_weighted, "rwc_indoor_weighted.rds")
saveRDS(sm_outdoor_weighted, "sm_outdoor_weighted.rds")
saveRDS(sm_indoor_weighted, "sm_indoor_weighted.rds")
```

```{r}
rwc_outdoor_weighted <- read_rds("rwc_outdoor_weighted.rds")
view(rwc_outdoor_weighted)
sm_outdoor_weighted <- read_rds("sm_outdoor_weighted.rds")
view(sm_outdoor_weighted)

epa_outdoor <- read_rds("epa_sensor_data.rds")
fc_outdoor <- read_rds("fc_sensor_data.rds")
smc_feb <- rwc_outdoor_weighted %>%
  rbind(
    sm_outdoor_weighted
  ) %>% rbind(epa_outdoor) %>% rbind(fc_outdoor) %>% rbind(rwc_indoor_weighted) %>% rbind(sm_indoor_weighted)


saveRDS(smc_feb, "smc_PM25_feb.rds")
```

```{r load-belmont-data-from-Awoe}
belmont <- read_rds("feb_daily_belmont_data.rds")

names(belmont)

smc_feb <- smc_feb %>%
  rbind(belmont)

saveRDS(smc_feb, "smc_PM25_feb.rds")
```

```{r create-popup-markers-for-mapview}
rwc_boundary %>%
  st_centroid()

smc_boundary %>%
  st_centroid()

saveRDS(smc_boundaries, "smc_boundaries.rds")

smc_boundaries <- places("CA", cb = T) %>%
  filter(NAME %in% c("Foster City", "Redwood City", "East Palo Alto", "Belmont", "San Mateo")) %>%
  mutate(
    NAME = if_else(NAME == "San Mateo", "San Mateo City", NAME)
  )
```

# Population Analysis

Indoor AQI is further coupled with census data to conduct population equity analysis. However remember in mind the representative indoor air quality data is calculated based on the closest sensor in the area. Indoor air quality varies significantly according to the ventilation of the house, cooking facilities, air purifiers, and etc. This may not serve as an accurate representation. The following analysis serves as a preliminary inspection.

```{r download-smc-cbgs}
library(tigris)
smc_cbgs <- block_groups("CA", county = "San Mateo County", cb = T)

saveRDS(smc_cbgs, "smc_cbgs.rds")
```

```{r read-smc-cbgs, eval=TRUE}
smc_cbgs <- read_rds("smc_cbgs.rds")
```

```{r}
library(leaflet)
leaflet() %>% 
  addProviderTiles(provider = providers$CartoDB.Positron) %>% 
  addPolygons(
    data = rwc_cbgs %>% st_transform(4326),
    fill = F
  ) %>%
  addPolygons(data = rwc_boundary, color = "Red")
```

```{r}
rwc_cbgs <- smc_cbgs %>%
  .[rwc_boundary, ]

sm_cbgs <- smc_cbgs %>%
  .[sm_boundary, ]
```

```{r rwc-cbg-pm25-1week}
rwc_pm25_in_voronoi_cbg <-
  rwc_pm25_in_voronoi %>% 
  st_intersection(rwc_cbgs %>% st_transform(4326)) %>% 
  st_make_valid() %>% 
  mutate(
    area = st_area(.) %>% as.numeric()
  ) %>% 
  st_drop_geometry() %>%
  mutate(pm2.5_1week = as.numeric(pm2.5_1week)) %>%
  group_by(GEOID) %>% 
  summarize(
    PM25 = weighted.mean(pm2.5_1week, area, na.rm = T)
  ) %>% 
  left_join(rwc_cbgs %>% dplyr::select(GEOID)) %>% 
  filter(PM25 != "NaN") %>%
  st_as_sf()


mapview(rwc_pm25_in_voronoi_cbg, zcol = "PM25")
mapview(rwc_boundary)
```

```{r sm-cbg-pm25-1week}
sm_pm25_in_voronoi_cbg <-
  sm_pm25_in_voronoi %>% 
  st_intersection(sm_cbgs %>% st_transform(4326)) %>% 
  st_make_valid() %>% 
  mutate(
    area = st_area(.) %>% as.numeric()
  ) %>% 
  st_drop_geometry() %>%
  mutate(pm2.5_1week = as.numeric(pm2.5_1week)) %>%
  group_by(GEOID) %>% 
  summarize(
    PM25 = weighted.mean(pm2.5_1week, area, na.rm = T)
  ) %>% 
  left_join(sm_cbgs %>% dplyr::select(GEOID)) %>% 
  filter(PM25 != "NaN") %>%
  st_as_sf()


mapview(sm_pm25_in_voronoi_cbg, zcol = "PM25")
mapview(sm_boundary)
```

```{r}
saveRDS(rwc_pm25_in_voronoi_cbg, "rwc_pm25_in_voronoi_cbg.rds")
saveRDS(sm_pm25_in_voronoi_cbg, "sm_pm25_in_voronoi_cbg.rds")
```

```{r}
smc_sensors_clean <- smc_sensors %>% 
  filter(
    !is.na(pm2.5_1week),
    !is.na(humidity)
  ) %>% 
  mutate(
    PM25 = 0.524*as.numeric(pm2.5_1week) - 0.0852*as.numeric(humidity) + 5.72,
    AQI = case_when(
      PM25 <= 12 ~ 
        paste(round(50/12*PM25), "Good"),
      PM25 <= 35.4 ~ 
        paste(round((100-51)/(35.4-12)*(PM25 - 12) + 51), "Moderate"),
      PM25 <= 55.4 ~
        paste(round((150-101)/(55.4-35.4)*(PM25 - 35.4) + 101), "Moderately Unhealthy"),
      PM25 <= 150.4 ~
        paste(round((200-151)/(150.4-55.4)*(PM25 - 55.4) + 151), "Unhealthy"),
      PM25 <= 250.4 ~
        paste(round((300-201)/(250.4-150.4)*(PM25 - 150.4) + 201), "Very Unhealthy"),
      TRUE ~ 
        paste(round((500-301)/(500.4-250.5)*(PM25 - 250.5) + 301), "Hazardous")
    )
  ) %>% 
  separate(
    AQI,
    into = c("AQI","AQI_Cat"),
    sep = " ",
    extra = "merge"
  ) %>% 
  mutate(
    AQI = as.numeric(AQI),
    AQI_Cat = AQI_Cat %>% factor(levels = c("Good", "Moderate","Moderately Unhealthy","Unhealthy","Very Unhealthy","Hazardous"))
  )
```

```{r}
aqi_pal <- colorFactor(
  palette = "RdYlGn",
  reverse = T,
  domain = smc_sensors_clean$AQI_Cat
)

smc_sensors_clean %>% 
  filter(location_type == "outside") %>% 
  leaflet() %>% 
  addProviderTiles(provider = providers$CartoDB.Positron) %>% 
  addCircleMarkers(
    color = ~aqi_pal(AQI_Cat),
    label = ~AQI_Cat,
    radius = 5,
    opacity = 0.75
  ) %>% 
  addLegend(
    pal = aqi_pal,
    values = ~AQI_Cat
  )
```

```{r sm-county-all-sensors-voronoi}
smc_in_voronoi <-
  smc_sensors_clean %>%
  filter(location_type == "inside") %>% 
  st_union() %>% 
  st_voronoi() %>% 
  st_cast() %>% 
  st_as_sf() %>% 
  st_intersection(.,st_union(smc_boundary %>% st_transform(4326))) %>% 
  st_join(smc_sensors_clean %>% filter(location_type == "inside"))

smc_out_voronoi <-
  smc_sensors_clean %>%
  filter(location_type == "outside") %>% 
  st_union() %>% 
  st_voronoi() %>% 
  st_cast() %>% 
  st_as_sf() %>% 
  st_intersection(.,st_union(smc_boundary %>% st_transform(4326))) %>% 
  st_join(smc_sensors_clean %>% filter(location_type == "outside"))

mapview(smc_in_voronoi, zcol = "AQI")
mapview(smc_out_voronoi, zol= "AQI")

smc_in_voronoi_cbg <-
  smc_in_voronoi %>% 
  st_intersection(smc_cbgs %>% st_transform(4326)) %>% 
  st_make_valid() %>% 
  mutate(
    area = st_area(.) %>% as.numeric()
  ) %>% 
  st_drop_geometry() %>%
  mutate(pm2.5_1week = as.numeric(pm2.5_1week)) %>%
  group_by(GEOID) %>% 
  summarize(
    AQI = weighted.mean(AQI, area, na.rm = T)
  ) %>% 
  left_join(smc_cbgs %>% dplyr::select(GEOID)) %>% 
  filter(AQI != "NaN") %>%
  st_as_sf()

smc_out_voronoi_cbg <-
  smc_out_voronoi %>% 
  st_intersection(smc_cbgs %>% st_transform(4326)) %>% 
  st_make_valid() %>% 
  mutate(
    area = st_area(.) %>% as.numeric()
  ) %>% 
  st_drop_geometry() %>%
  mutate(pm2.5_1week = as.numeric(pm2.5_1week)) %>%
  group_by(GEOID) %>% 
  summarize(
    AQI = weighted.mean(AQI, area, na.rm = T)
  ) %>% 
  left_join(smc_cbgs %>% dplyr::select(GEOID)) %>% 
  filter(AQI != "NaN") %>%
  st_as_sf()

mapview(smc_in_voronoi_cbg, zcol = "AQI")
mapview(smc_out_voronoi_cbg, zcol = "AQI")

saveRDS(smc_in_voronoi_cbg, "smc_in_voronoi_cbg_AQI.rds")
saveRDS(smc_out_voronoi_cbg, "smc_out_voronoi_cbg_AQI.rds")
```

```{r sm-county-voronoi-cbgs&blocks}
smc_in_voronoi_cbg <-
  smc_in_voronoi %>% 
  st_intersection(smc_cbgs %>% st_transform(4326)) %>% 
  st_make_valid() %>% 
  mutate(
    area = st_area(.) %>% as.numeric()
  ) %>% 
  st_drop_geometry() %>%
  mutate(pm2.5_1week = as.numeric(pm2.5_1week)) %>%
  group_by(GEOID) %>% 
  summarize(
    PM25 = weighted.mean(pm2.5_1week, area, na.rm = T)
  ) %>% 
  left_join(smc_cbgs %>% dplyr::select(GEOID)) %>% 
  filter(PM25 != "NaN") %>%
  st_as_sf()

saveRDS(smc_in_voronoi_cbg, "smc_in_voronoi_cbg_1week.rds")

smc_blocks <- blocks(state = "06", county = "081")

smc_in_voronoi_blocks <-
  smc_in_voronoi %>% 
  st_intersection(smc_blocks %>% st_transform(4326)) %>% 
  st_make_valid() %>% 
  mutate(
    area = st_area(.) %>% as.numeric()
  ) %>% 
  st_drop_geometry() %>%
  mutate(pm2.5_1week = as.numeric(pm2.5_1week)) %>%
  group_by(GEOID) %>% 
  summarize(
    PM25 = weighted.mean(pm2.5_1week, area, na.rm = T)
  ) %>% 
  left_join(smc_blocks %>% dplyr::select(GEOID)) %>% 
  filter(PM25 != "NaN") %>%
  st_as_sf()

saveRDS(smc_in_voronoi_blocks, "smc_in_voronoi_blocks_1week.rds")
```

```{r sm-county-1week-inside-map, eval = T}
smc_in_voronoi_cbg <- read_rds("smc_in_voronoi_cbg_1week.rds")

mapview(smc_in_voronoi_cbg, zcol = "PM25")
```

```{r census-metadata}
library(tidyverse)
library(censusapi)

Sys.setenv(CENSUS_KEY="c8aa67e4086b4b5ce3a8717f59faa9a28f611dab")

acs_vars_2019_5yr <- read_rds("acs_vars_2019_5yr.rds")
```

```{r}
names(smc_in_voronoi_cbg)
```

```{r}
names(smc_cbgs)

blocks_pop <-
  getCensus(
    name = "dec/pl",
    vintage = 2020,
    region = "block:*", 
    regionin = "state:06+county:081",
    vars = "P1_001N"
  ) %>% 
  transmute(
    GEOID20 = paste0(state,county,tract,block),
    pop = P1_001N
  ) %>%
  mutate(GEOID = substr(GEOID20, 1, 12)) %>%
  group_by(GEOID) %>%
  summarize(pop =sum(pop))
```

```{r pop-map}
smc_aqi_pop <- smc_in_voronoi_cbg %>%
  mutate(Location = "inside") %>%
  rbind(smc_out_voronoi_cbg %>% mutate(Location = "outside")) %>%
  mutate(
    AQI_level = case_when(
      AQI > 50 ~ "Bad", 
      AQI <= 50 & AQI > 30 ~ "Moderate",
      AQI <= 30 & AQI > 20 ~ "Good",
      AQI <= 20 ~ "Very good"
    ) %>% factor(levels = c("Bad", "Moderate", "Good", "Very good"))
  ) %>%
  left_join(
    blocks_pop
  ) %>% mutate(
    pop = if_else(is.na(pop), 0, pop)
  )
  

saveRDS(smc_aqi_pop, "smc_aqi_pop.rds")

smc_aqi_pop <- read_rds("smc_aqi_pop.rds")

aqi_pal <- colorNumeric(
  palette = "Purples",
  domain = smc_aqi_pop$AQI
)

leaflet() %>%
  addProviderTiles(provider = providers$CartoDB.Positron) %>%
  addPolygons(data = smc_aqi_pop %>% st_transform(4326),
              fillColor = ~aqi_pal(AQI),
              color = "white",
              opacity = 0.5,
              fillOpacity = 0.8,
              weight = 1,
              label = ~paste0(
                "Average weekly AQI ", round(AQI), " in CBG", GEOID, " with ", pop, " population"
              ))

mapview(smc_aqi_pop, zcol = "AQI") +
  mapview(smc_aqi_pop %>% st_centroid(), zcol = "pop", cex = "pop")

smc_aqi_pop <- read_rds("smc_aqi_pop.rds")

smc_pop_plot <- smc_aqi_pop %>%
  filter(Location == "inside") %>%
  group_by(AQI_level) %>%
  summarize(pop = sum(pop, na.rm = T))
saveRDS(smc_pop_plot, "smc_pop_plot.rds") 
```

```{r pop-plot}
smc_pop_plot <- read_rds("smc_pop_plot.rds")
smc_pop_plot %>% 
  ggplot() +
  geom_bar(
    aes(x = AQI_level,
        y = pop),
    stat = "identity"
  ) + 
  labs(
    x = "AQI level",
    y = "Population",
    title = "Population exposure"
  )
```

```{r smc-tenure-data-processing}
smc_pop_tenure_units <-
  getCensus(
    name = "acs/acs5",
    vintage = 2019,
    region = "block group:*", 
    regionin = "state:06+county:081",
    vars = "group(B25033)"
  ) %>% 
  mutate(
    cbg =
      paste0(state,county,tract,block_group)
  ) %>% 
  select(!c(GEO_ID,state,county,tract,block_group,NAME) & !ends_with(c("EA","MA","M"))) %>%
  pivot_longer(
    ends_with("E"),
    names_to = "variable",
    values_to = "estimate"
  ) %>%
  left_join(
    acs_vars_2019_5yr %>% 
      select(name, label), 
    by = c("variable" = "name")
  ) %>% 
  select(-variable) %>% 
  separate(
    label,
    into = c(NA,NA,"tenure","units_in_structure"),
    sep = "!!"
  ) 

smc_tenure <- 
  smc_pop_tenure_units %>%
  filter(
    is.na(units_in_structure),
    !is.na(tenure)
  ) %>%
  select(-units_in_structure)

smc_tenure <- smc_tenure %>%
  mutate(
    tenure = case_when(
      tenure == "Owner occupied:" ~ "Owner",
      tenure == "Renter occupied:" ~ "Renter"
    )
  ) %>%
  pivot_wider(
    names_from = "tenure",
    values_from = "estimate"
  )
```

```{r smc-tenure-plot}
plot(smc_in_voronoi_cbg$AQI)

smc_equity <-
  smc_in_voronoi_cbg %>%
  mutate(
    AQI_level = case_when(
      AQI > 50 ~ "Bad", 
      AQI <= 50 & AQI > 30 ~ "Moderate",
      AQI <= 30 & AQI > 20 ~ "Good",
      AQI <= 20 ~ "Very good"
    )
  ) %>%
  left_join(smc_tenure, by = c("GEOID" = "cbg"))

smc_tenure_plot <-
  smc_equity %>% #having geometry for each CBGs
  group_by(
    AQI_level
  ) %>%
  summarize(
    Owner = sum(Owner),
    Renter = sum(Renter)
  ) %>%
  st_drop_geometry() %>%
  pivot_longer(
    cols = c("Owner", "Renter"),
    names_to = "tenure",
    values_to = "estimate"
  )

saveRDS(smc_tenure_plot, "smc_tenure_plot.rds")
```

```{r tenure-plot}
smc_tenure_plot <- read_rds("smc_tenure_plot.rds")

smc_tenure_plot %>%
  ggplot() +
  geom_bar(
    aes(
      x = AQI_level %>% factor(levels = rev(c("Very good", "Good", "Moderate", "Bad"))),
      y = estimate,
      fill = tenure
    ),
    stat = "identity",
    position = "fill"
  ) + 
  labs(
    x = "AQI level",
    y = "Population",
    title = "San Mateo County AQI level by tenure",
    fill = "Household tenure"
  ) + coord_flip()
```

```{r smc-income-data-processing}
smc_income <-
  getCensus(
    name = "acs/acs5",
    vintage = 2019,
    region = "block group:*",
    regionin = "state:06+county:081",
    vars = "group(B19001)"
  ) %>%
  mutate(
    cbg =
      paste0(state,county,tract,block_group)
  ) %>%
  select(!c(GEO_ID,state,NAME) & !ends_with(c("EA","MA","M"))) %>%
  pivot_longer(
    ends_with("E"),
    names_to = "name",
    values_to = "estimate"
  ) %>%
  left_join(
    acs_vars_2019_5yr %>% 
      select(name, label)
  ) %>% 
  select(-name) %>% 
  separate(
    label,
    into = c(NA,NA,"income"),
    sep = "!!"
  ) %>% 
  filter(!is.na(income)) %>%
  select(cbg, estimate, income)

saveRDS(smc_income, "smc_income.rds")
```

```{r smc-income-plot-processing}
smc_income_plot <-
  smc_in_voronoi_cbg %>%
  mutate(
    AQI_level = case_when(
      AQI > 50 ~ "Bad", 
      AQI <= 50 & AQI > 30 ~ "Moderate",
      AQI <= 30 & AQI > 20 ~ "Good",
      AQI <= 20 ~ "Very good"
    )
  ) %>%
  left_join(smc_income, by = c("GEOID" = "cbg"))

smc_income_plot <- smc_income_plot %>% #having geometry for each CBGs
  group_by(
    AQI_level, income
  ) %>%
  summarize(estimate = sum(estimate)) %>%
  st_drop_geometry()

saveRDS(smc_income_plot, "smc_income_plot.rds")

unique(smc_income$income)
```

```{r smc-income-plot}
smc_income_plot<-read_rds("smc_income_plot.rds")
smc_income_plot %>% #having geometry for each CBGs
  ggplot() +
  geom_bar(
    aes(
      x = AQI_level %>% factor(levels = rev(c("Very good", "Good", "Moderate", "Bad"))),
      y = estimate,
      fill = income %>% factor(levels = rev(unique(smc_income_plot$income)))
    ),
    stat = "identity",
    position = "fill"
  ) + 
  labs(
    x = "AQI level",
    y = "Population",
    title = "San Mateo County AQI level by income",
    fill = "Income level"
  ) + coord_flip()
```

```{r smc-race-data-processing}
dec_vars_2020 <-
  listCensusMetadata(
    name = "2020/dec/pl",
    type = "variables"
  )

smc_race <- 
  getCensus(
    name = "dec/pl",
    vintage = 2020,
    region = "block:*",
    regionin = "state:06+county:081",
    vars = "group(P1)"
  ) %>% 
  mutate(
    block =
      paste0(state,county,tract,block)
  ) %>% 
  select(!c(GEO_ID,state,county,tract,NAME) & !ends_with(c("NA"))) %>% 
  pivot_longer(
    ends_with("N"),
    names_to = "name",
    values_to = "estimate"
  ) %>%
  left_join(
    dec_vars_2020 %>% 
      select(name, label)
  ) %>% 
  select(-name) %>% 
  separate(
    label,
    into = c(NA,NA,"category1","category2"),
    sep = "!!")

smc_race_plot <- 
  smc_race %>%
  filter(!is.na(category1)) %>%
  mutate(
    race = case_when(
      category1 == "Population of one race:" ~category2,
      category1 == "Population of two or more races:" ~ "More than one race"
    )
  ) %>% filter(!is.na(race)) %>%
  select(-category1, -category2) %>%
  group_by(block,race) %>%
  summarize(
    estimate = sum(estimate)
  ) %>%
  mutate(
    cbg = substr(block, 1,12)
  ) %>%
  group_by(cbg,race) %>%
  summarize(estimate = sum(estimate))

smc_race_plot <-
  smc_in_voronoi_cbg %>%
  mutate(
    AQI_level = case_when(
      AQI > 50 ~ "Bad", 
      AQI <= 50 & AQI > 30 ~ "Moderate",
      AQI <= 30 & AQI > 20 ~ "Good",
      AQI <= 20 ~ "Very good"
    )
  ) %>%
  left_join(smc_race_plot, by = c("GEOID" = "cbg")) %>%
  st_drop_geometry() 
saveRDS(smc_race_plot, "smc_race_plot.rds")
```

There are significant amount of Asian people living in bad air quality area.
```{r smc-race-plotting}
smc_race_plot <- read_rds("smc_race_plot.rds")

smc_race_plot %>% #having geometry for each CBGs
  ggplot() +
  geom_bar(
    aes(
      x = AQI_level %>% factor(levels = rev(c("Very good", "Good", "Moderate", "Bad"))),
      y = estimate,
      fill = race %>% factor(levels = rev(unique(smc_race_plot$race)))
    ),
    stat = "identity",
    position = "fill"
  ) + 
  labs(
    x = "AQI level",
    y = "Population",
    title = "San Mateo County AQI level by race",
    fill = "race level"
  ) + coord_flip()
```

```{r}
names(test)

test <- smc_race_plot %>%
  rename(category = race)

saveRDS(test, "dash_race_plot.rds")
```

