---
title: "Sustainability Analysis: GHG Emission"
author: "Spencer Zhang"
date: "2/7/2022"
output: bookdown::html_document2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=T}
knitr::opts_chunk$set(echo = T, warning = F, message = F, eval = F)
options(tigris_use_cache = TRUE)
```

```{r library}
library(tidyverse)
library(sf)
library(tigris)
library(censusapi)
library(mapview)
library(leaflet)
library(mapboxapi)
library(readr)
library(jsonlite)
```

# Vehicle Emission
## Commute job counts from LODES
Origin-destination job counts by block to block level

ZIP code, blocks data preparation
```{r zip-94108, eval = F}
zctas <- zctas() 
#Download a Zip Code Tabulation Area (ZCTA) shapefile into R

#downtown Palo Alto residential area
zip <- zctas %>%
  filter(GEOID10 == "94108") #single line information of 94108

blocks <- blocks("CA") #all block information in CA

zip_blocks <- blocks %>%
  st_centroid() %>%
  .[zip, ] #all blocks in zip=94108
saveRDS(zip, "zip_info_94108.rds")
saveRDS(zip_blocks, "zip_block_info_94108.rds")
```

```{r read-zip}
zip <- read_rds("zip_info_94108.rds")
zip_blocks <- read_rds("zip_block_info_94108.rds")
zip_cbgs <- 
  block_groups(state = "CA", county = "San Francisco") %>% 
  st_centroid() %>%
  .[zip,] %>% 
  st_drop_geometry() %>% 
  left_join(block_groups(state = "CA", county = "San Francisco") %>% select(GEOID)) %>% 
  st_as_sf()
```

LODES OD data download and save locally
```{r LODES-commute-od-data, eval = F}
full_ct_od <- 2013:2019 %>%
  map_dfr(function(year){

    print(year)

    temp <- read_csv(paste0("/Volumes/GoogleDrive/Shared drives/SFBI/Data Library/LODES/ca_od_main_JT01_", year, ".csv.gz")) %>%
      filter(
        h_geocode %in% zip_blocks$GEOID10 |
          w_geocode %in% zip_blocks$GEOID10
      ) %>%
      mutate(year = year)

    saveRDS(temp, paste0("94108_od_", year, ".rds"))

    return(temp)

  }) #5 years data

saveRDS(full_ct_od, "94108_od_full.rds")
```

read LODES OD local file
```{r read-commute-od-data}
full_ct_od <- readRDS("94108_od_full.rds")
#full job-to-job OD data where either work location or home location is in 94108
#block-to-block level
```

categorize OD data entries by whether or not crossed the ZIP code boundary 
"within": home and workplace are within ZIPcode
"outbound": home in ZIP, workplace outside of ZIP
"inbound": home outside of ZIP, workplace in ZIP
```{r job-od-categorization}
#separate the jobs into inbound, outbound, and within the same zip code
zip_od <- full_ct_od %>%
  select(-createdate) %>%
  mutate(
    direction = case_when(
      h_geocode %in% zip_blocks$GEOID10 &
        w_geocode %in% zip_blocks$GEOID10 ~ "within",
      h_geocode %in% zip_blocks$GEOID10 ~ "outbound",
      w_geocode %in% zip_blocks$GEOID10 ~ "inbound"
    )
  )

table(zip_od$direction)
# inbound outbound   within 
# 162086    39853     2896 
```

```{r add-inbound/outbound-to-LODES, eval = F}
full_ct_od_cross <- full_ct_od %>%
  select(-createdate) %>%
  filter(!(
    h_geocode %in% zip_blocks$GEOID10 &
      w_geocode %in% zip_blocks$GEOID10
  )) %>% #not include people who travel within their zip code
  mutate(
    direction = ifelse(
      h_geocode %in% zip_blocks$GEOID10,
      "outbound",
      "inbound"
    )
  ) #commute data where people travel out or into 94108

table(full_ct_od_cross$direction)
```

## Routing from home to workplace 
The original OD data is on block-to-block level. To simplify the routing, we group data into **tract to ZIP code** level for later routing with mapbox services.
```{r filter-out-routing-directions, eval = F}
full_ct_od_routing <- full_ct_od_cross %>%
  mutate(
    origin = ifelse(
      direction == "inbound",
      h_geocode,
      w_geocode
    ),
    cbg = origin %>% substr(1,12),
    tract = origin %>% substr(1,11)
  ) %>%
  filter(!duplicated(tract))
# Question: why are we filtering out the duplicated??
# Answer: to do routing only once for the same origin
```

Routing for ODs that crosses ZIP code boundary.
Remember there are ODs all within the ZIP code boundary not counted here.
```{r routing-mapbox, eval = F}
# ca_tracts <- tracts("CA") #census acs5/1 only to tract levels
#only decennial data offers blockgroup levels
# 
# #routing origins (all inbound/outbound locations)
od_origin <-
  full_ct_od_routing %>%
  select(tract) %>%
  left_join(ca_tracts %>% select(tract = GEOID)) %>%
  st_as_sf() %>%
  st_centroid() %>%
  st_coordinates() #add geometry information into the od_origins for later routing

view(head(full_ct_od_routing)) #contains destination(outside of 94108) tract ids
view(head(od_origin)) #centroid coordinate(X,Y) for  


# nrow(od_origin) #7044

#routing destination (94108 only)
od_destination <-
  zip %>%
  st_centroid() %>%
  st_coordinates()

#routing using mapbox
od_route <-
  1:ceiling(nrow(od_origin)/1000) %>%
  map_dfr(function(y){

    print(y)

    temp <-
      (y * 1000 - 999) : pmin(y * 1000, nrow(od_origin)) %>%
      map_dfr(function(x){
        tryCatch(
          mb_directions(
            origin = od_origin[x, ],
            destination = od_destination,
            profile = "driving-traffic"
          ) %>%
            mutate(id = x),
          error = function(e){
            data.frame(id = x)
          }
        )
      }) %>%
      st_as_sf()

      saveRDS(temp, paste0("94108_od_routing_",y,".rds"))

      return(temp)
  })
saveRDS(od_route, "94108_od_routing_full.rds")

view(head(od_route))
```

```{r read-commute-od-routed-duration-time, eval = F}
od_route <- readRDS("94108_od_routing_full.rds")
nrow(full_ct_od_routing)

od_route %>%
  filter(is.na(duration)) #2 rows
```

```{r combine-routed-results-with-jobs, eval = F}
full_od_routed <- full_ct_od_routing %>%
  cbind(od_route)

view(head(full_od_routed))
view(head(full_ct_od_cross))
full_od_routed %>%
  filter(is.na(duration)) #2 rows
names(full_od_routed)

sum(is.na(full_od_final$duration))

view(head(full_od_final))

full_od_final <- full_ct_od_cross %>%
  mutate(
    origin = ifelse(
      direction == "inbound",
      h_geocode,
      w_geocode
    ),
    tract = substr(origin, 1, 11)
  ) %>%
  left_join(
    full_od_routed %>%
      select(tract, duration, distance)
  ) %>%
  mutate(
    visits = S000 * 261 #TODO: the number of visits per year for commute patterns
  )
saveRDS(full_od_final,"94108_commute_od_routed_2013_2019.rds")
```

```{r read-commute-routed-2013-2019}
full_od_final <- read_rds("94108_commute_od_routed_2013_2019.rds")

view(head(full_od_final))
```

```{r}
# ODs with routing results
nrow(full_od_final)
# 201939

# original full ODs dataset
nrow(full_ct_od)
# 204835
# 3k entries more, from the OD all within zip code boundary trips
```

### routing for within ODs
to address the nearly 3k block-to-block that are actually within the same ZIPcode
```{r filter-within-OD-trips}
within_routing <- zip_od %>%
  filter(direction == "within") %>%
  group_by(
    w_geocode, h_geocode
  ) %>%
  summarize(n = n(), S000 = sum(S000))

view(head(within_routing))

table(within_routing$n)
#   1   2   3   4   5   6   7 
# 759 337 161 100  53  35  15 
```

To simplify, we route within-OD trips by workplace location to centroid of ZIPcode
```{r routing-for-within-OD-trips}
od_within_origin <-
  within_routing %>%
  select(w_geocode) %>%
  left_join(zip_blocks %>% select(w_geocode = GEOID10)) %>%
  st_as_sf() %>%
  st_centroid() %>%
  st_coordinates() #add geometry information into the od_origins for later routing

# nrow(od_within_origin) #1460

#routing destination (94108 only)
od_destination <-
  zip %>%
  st_centroid() %>%
  st_coordinates()

#routing using mapbox
od_within_route <-
  1:ceiling(nrow(od_within_origin)/100) %>%
  map_dfr(function(y){

    print(y)

    temp <-
      (y * 100 - 99) : pmin(y * 100, nrow(od_within_origin)) %>%
      map_dfr(function(x){
        tryCatch(
          mb_directions(
            origin = od_within_origin[x, ],
            destination = od_destination,
            profile = "driving-traffic"
          ) %>%
            mutate(id = x),
          error = function(e){
            data.frame(id = x)
          }
        )
      }) %>%
      st_as_sf()

      saveRDS(temp, paste0("94108_od_within_routing_",y,".rds"))

      return(temp)
  })
saveRDS(od_within_route, "94108_od_within_routing_full.rds")

view(head(od_within_route))
```

```{r}
within_routed <- within_routing %>%
  cbind(od_within_route) %>%
  select(-n, -S000, -id)
```

```{r full-OD-routed-data, eval = F}
full_routed <-
  zip_od %>%
  filter(direction == "within") %>%
  left_join(
    within_routed, by = c("w_geocode", "h_geocode")
  ) %>%
  mutate(
    visits = S000*261,
    origin = h_geocode,
    tract = substr(h_geocode, 1, 11)
  ) %>% select(-geometry) %>%
  rbind(
    full_od_final
  )

saveRDS(full_routed, "full_routed_94108.rds")
```

### duration, distance, visits of FULL OD in zip94108
```{r}
full_routed <- read_rds("full_routed_94108.rds")
```

## Grab population data from ACS5
```{r pop-data}
#convert distance and visits in full_od_final into VMT
#visits are "per job" so combine with census data to calculate "per vehicle"
Sys.setenv(CENSUS_KEY="c8aa67e4086b4b5ce3a8717f59faa9a28f611dab")

ca_cbgs_pop <-
  counties("CA", cb = T, progress_bar = F) %>%
  pull(COUNTYFP) %>%
  map_dfr(function(x){
    getCensus(
      name = "acs/acs5",
      vintage = 2019,
      region = "block group:*",
      regionin = paste0("state:06+county:", x),
      vars = "B01001_001E"
    )
  }) %>%
  transmute(
    census_block_group =
      paste0(state,county,tract,block_group),
    pop = B01001_001E
  )
#names(ca_cbgs_pop) #"census_block_group" "pop"

test <-
  2013:2019 %>%
  map_dfr(function(year){
    counties("CA", cb = T, progress_bar = F) %>%
    pull(COUNTYFP) %>%
    map_dfr(function(x){
      getCensus(
        name = "acs/acs5",
        vintage = year,
        region = "block group:*",
        regionin = paste0("state:06+county:", x),
        vars = "B01001_001E"
      )
    }) %>%
    transmute(
      census_block_group =
        paste0(state,county,tract,block_group),
      pop = B01001_001E,
      year = year
    )
  })
view(head(ca_cbgs_pop))

saveRDS(test, "ca_pop_2013_2019.rds")
ca_cbgs_pop <- readRDS("ca_pop_2013_2019.rds")

origin_cbgs_pop <-
  full_routed %>%
  select(-starts_with("S")) %>%
  mutate(
    h_cbgs = substr(h_geocode, 1, 12)
  ) %>%
  left_join(
    ca_cbgs_pop,
    by = c("h_cbgs" = "census_block_group", "year" = "year")
  )

saveRDS(origin_cbgs_pop, "origin_cbgs_pop_94108.rds")

origin_cbgs_pop <- read_rds("origin_cbgs_pop_94108.rds")
```

There are three census block groups that have no population documented, excluded from this analysis. They are 060379304011, 060378002043, 060378002043.
```{r eval = F}
sum(is.na(origin_cbgs_pop$pop)) #3
sum(is.na(origin_cbgs_pop$duration)) #4

#TODO: find out where are these places and why they have no routing as well
origin_cbgs_pop %>% filter(is.na(pop))
#060379304011
#060378002043
#060378002043
```

## Create travel mode for later turning job counts into vehicle counts
Since not every commuter travel by driving alone, we need the travel mode information for all the related tracts to convert our job counts, duration, travel time into actual vehicle miles traveled.
We assume the travel mode is constant from 2013 to 2019.
```{r travel-mode-by-time}
acs_vars_2019_5yr <- read_rds("acs_vars_2019_5yr.rds")

travel_time_mode <-
  counties("CA", cb = T, progress_bar = F) %>%
  pull(COUNTYFP) %>%
  map_dfr(function(x){
    getCensus(
      name = "acs/acs5",
      vintage = 2019,
      region = "block group:*",
      regionin = paste0("state:06+county:", x),
      vars = "group(B08134)" #Mode to work by travel time to work
    )
  }) %>%
  mutate(
    cbg =
      paste0(state,county,tract,block_group)
  ) %>%
  filter(cbg %in% origin_cbgs_pop$h_cbgs) %>%
  select(!c(GEO_ID,state,county,tract,block_group,NAME) & !ends_with(c("EA","MA","M"))) %>%
  pivot_longer(
    ends_with("E"),
    names_to = "variable",
    values_to = "estimate"
  ) %>%
  left_join(
    acs_vars_2019_5yr %>%
      select(name, label),
    by = c("variable" = "name")
  ) %>%
  select(-variable) %>%
  separate(
    label,
    into = c(NA, NA, "total", "mode", "carpool", "time"),
    sep = "!!"
  ) %>%
  mutate(
    mode = case_when(
      total %in% c(
        "Less than 10 minutes",
        "10 to 14 minutes",
        "15 to 19 minutes",
        "20 to 24 minutes",
        "25 to 29 minutes",
        "30 to 34 minutes",
        "35 to 44 minutes",
        "45 to 59 minutes",
        "60 or more minutes"
      ) ~ "Total",
      mode == "Drove alone:" ~ mode,
      carpool %in% c(
        "In 2-person carpool:",
        "In 3-or-more-person carpool:"
      ) ~ carpool
    ),
    time = case_when(
      mode == "Total" ~ total,
      mode == "Drove alone:" ~ carpool,
      mode == carpool ~ time
    )
  ) %>%
  filter(!is.na(time)) %>%
  select(-total, -carpool) %>%
  pivot_wider(
    names_from = mode,
    values_from = estimate
  ) %>%
  mutate(
    perc_veh1 = `Drove alone:`/Total,
    perc_veh2 = `In 2-person carpool:`/Total,
    perc_veh3 = `In 3-or-more-person carpool:`/Total
  )

saveRDS(travel_time_mode, "94108_travel_time_mode.rds")

travel_time_mode <- read_rds("94108_travel_time_mode.rds")
```

### combine travel mode with job count trips -> get vehicle counts
```{r ct-trips-construct, eval = F}
ct_trips <-
  origin_cbgs_pop %>% 
  mutate(
    time = case_when(
      duration < 10 ~ "Less than 10 minutes",
      duration < 15 ~ "10 to 14 minutes",
      duration < 20 ~ "15 to 19 minutes",
      duration < 25 ~ "20 to 24 minutes",
      duration < 30 ~ "25 to 29 minutes",
      duration < 35 ~ "30 to 34 minutes",
      duration < 45 ~ "35 to 44 minutes",
      duration < 60 ~ "45 to 59 minutes",
      TRUE ~ "60 or more minutes"
    )
  ) %>% 
  left_join(
    travel_time_mode %>% 
      select(
        cbg,
        time,
        perc_veh1,
        perc_veh2,
        perc_veh3
      ),
    by = c("h_cbgs" = "cbg", "time")
  ) %>% 
  mutate(
    vehicles = 
      visits * perc_veh1 + 
      visits * perc_veh2 / 2 +
      visits * perc_veh3 / 3,
    vmt = vehicles * distance * 2
  )
saveRDS(ct_trips, "94108_trips.rds")
```

```{r visits-totalVMT-by-year-origin-geocode15, eval = F}
ct_trips <- read_rds("94108_trips.rds")

ct_visits <- ct_trips %>%
  group_by(tract, year) %>%
  summarize(
    visits = sum(visits),
    duration = mean(duration),
    distance = mean(distance),
    vmt = sum(vmt)
  ) %>%
  st_drop_geometry() %>% 
  left_join(tracts(state = "CA") %>% select(tract = GEOID)) %>%
  st_as_sf()

saveRDS(ct_visits, "94108_trips_vmt_by_year_tract.rds")
```

```{r load-vmt-data}
ct_trips <- read_rds("94108_trips.rds")
ct_visits <- read_rds("94108_trips_vmt_by_year_tract.rds")
```

```{r vehicle-routing-plot}
#TODO: fix this chunk
visits_pal <- colorNumeric(
  palette = "Reds",
  domain = ct_visits %>% 
    filter(year == 2019) %>%
    arrange(desc(visits)) %>% 
    pull(visits) 
    # %>% .[-c(1:6)]
)

#TODO: this plot is way too slow 
leaflet() %>% 
  addMapboxTiles(
    style_id = "light-v9",
    username = "mapbox"
  ) %>% 
  addPolygons(
    data = zip_cbgs %>% st_transform(4326),
    fill = F
  ) %>% 
  addPolygons(
    data = ct_visits %>% st_transform(4326),
    fillColor = ~visits_pal(visits),
    color = "red",
    weight = 1,
    fillOpacity = 0.75,
    label = ~visits
  ) %>% 
  addLegend(
    data = ct_visits,
    pal = visits_pal,
    values = ct_visits %>% st_transform(4326) %>% 
      filter(year == 2019) %>%
      arrange(desc(visits)) %>% 
      pull(visits), 
    # %>% .[-c(1:6)],
    title = "Commute visits to 94108"
  )
```

##convert VMTs to GHG emission
```{r emfac-factors}
emfac <- 
  read_csv("EMFAC2021-EI-202xClass-SanFranciscoBayArea-2021-Summer.csv", skip = 8) %>% 
  transmute(
    Category = `Vehicle Category`,
    Fuel_Type = Fuel,
    Percent_Trips = Trips/sum(Trips),
    Percent_Miles = `Total VMT`/sum(`Total VMT`),
    `MTCO2_Running_Exhaust` = CO2_RUNEX/`Total VMT`,
    `MTCO2_Start_Exhaust` = CO2_STREX/Trips
  )
```

```{r GHG-by-year-total,eval = F}
ct_trips_ghg_by_year <-
  2013:2019 %>%
  map_dfr(function(y){
    emfac %>% 
  mutate(
    trips = Percent_Trips * sum(ct_trips %>% filter(year == y) %>% .$visits, na.rm = T),
    vmt = Percent_Miles * sum(ct_trips %>% filter(year == y) %>% .$vmt, na.rm = T),
    ghg = vmt*MTCO2_Running_Exhaust + trips*MTCO2_Start_Exhaust*2,
    year = y
  )
  })

ct_trips_ghg_annual <-
  ct_trips_ghg_by_year %>%
  group_by(year) %>%
  summarize(trips = sum(trips),
            vmt = sum(vmt),
            ghg = sum(ghg))

saveRDS(ct_trips_ghg_annual, "94108_vmt_ghg_total.rds")
```


# Building Emission
```{r PGE-elec-emission-factor-by-year}
#Benchmarking Greenhouse Gas Emissions for Delivered Electricity 
#UNIT: (Pounds of CO2 per MWh)

pge_elec_emissions_factor <-
  data.frame(
    year = c(2013:2019),
    factor = c(427,435,405,294,210,206,2.68)
  )

pge_elec_emissions_factor %>% 
  ggplot() +
  geom_line(
    aes(
      x = year,
      y = factor
    )
  ) +
  labs(
    x = "Year",
    y = "Pounds of CO2 per MHh",
    title = "PG&E electricity emissions rate"
  )
```

```{r PGE-gas-emission-factor}
#assume it's constant over years
#UNIT: (pounds of CO2 per million BTU)
pge_gas_emissions_factor <- 117
```

```{r pge-by-year-data, eval = F}
pge_data <- 
  2013:2019 %>% 
  map_dfr(function(yr){
    
    factor <- 
      pge_elec_emissions_factor %>% 
      filter(year == yr) %>% 
      pull(factor)
    
    1:4 %>% 
      map_dfr(function(quarter){
        
        c("Electric","Gas") %>% 
          map_dfr(function(type){
            
            filename <- 
              paste0(
                "/Volumes/GoogleDrive/Shared drives/SFBI/Data Library/PG&E/PGE_",
                yr,
                "_Q",
                quarter,
                "_",
                type,
                "UsageByZip.csv"
              )
            
            temp <- read_csv(filename)
            
            if(yr == 2017 & quarter == 4) {
              temp <- 
                temp %>% 
                filter(MONTH != 9)
            }
            
            temp <-
              temp %>% 
              rename_all(toupper) %>% 
              mutate(
                TOTALKBTU = ifelse(
                  substr(CUSTOMERCLASS,1,1) == "E",
                  TOTALKWH * 3.412,
                  TOTALTHM * 99.976
                ),
                TOTALTCO2E = ifelse(
                  substr(CUSTOMERCLASS,1,1) == "E",
                  TOTALKWH/1000 * factor * 0.000453592,
                  TOTALTHM * 0.00531
                )
              ) %>% 
              select(
                ZIPCODE,
                YEAR,
                MONTH,
                CUSTOMERCLASS,
                TOTALKBTU,
                TOTALTCO2E,
                TOTALCUSTOMERS
              )
            
          })
        
      })
    
  })

saveRDS(pge_data, "pge_data_2013_2019.rds")
```

```{r building-in-94108, eval=F}
pge_data <- read_rds("pge_data_2013_2019.rds")

bldg_pge_by_year <- pge_data %>% 
  filter(ZIPCODE == 94108) %>% 
  filter(CUSTOMERCLASS %in% c(
    "Elec- Commercial",
    "Elec- Residential",
    "Gas- Commercial",
    "Gas- Residential"
  )) %>% 
  mutate(
    ENERGYTYPE = substr(CUSTOMERCLASS,1,1),
    BUILDINGTYPE = case_when(
      ENERGYTYPE == "E" ~ substr(CUSTOMERCLASS,7,7),
      ENERGYTYPE == "G" ~ substr(CUSTOMERCLASS,6,6)
    )
  ) %>% 
  group_by(ZIPCODE, ENERGYTYPE, BUILDINGTYPE, YEAR) %>%
  summarize(
    TOTALKBTU = sum(TOTALKBTU, na.rm=T),
    TOTALTCO2E = sum(TOTALTCO2E, na.rm=T),
    TOTALCUSTOMERS = mean(TOTALCUSTOMERS, na.rm=T)
  ) 
# %>%
#   group_by(ENERGYTYPE, BUILDINGTYPE, YEAR) %>% #the second group by doesnt make any sense because we are just dealing with one ZIP code here, but it doesnt hurt to keep here
#   summarize(across(
#     c(TOTALKBTU,TOTALTCO2E, TOTALCUSTOMERS),
#     ~sum(.,na.rm=T)
#   ))

saveRDS(bldg_pge_by_year, "94108_bldg_pge_2013_2019.rds")
```

```{r read-bldg_pge_by_year}
bldg_pge_by_year <- read_rds("94108_bldg_pge_2013_2019.rds")
```

```{r}
#TODO
ggplot(
  bldg_pge_by_year, 
  aes(
    x = as.factor(YEAR), 
    y = TOTALKBTU/1000000
  )
) + 
  geom_bar(stat = "identity", aes(fill = ENERGYTYPE), position = "dodge") + 
  labs(x = "Year", y = "GBTU", title = "94108 Annual Energy Usage, 2013 to 2019") + 
  scale_fill_discrete(name="Energy Type",labels = c("Electricity","Gas"))
```

## building energy consumption by resident
```{r resident-data, eval =F}
#get population data in ZIP=94108 2013-2019
#DATA: acs5
# names(ca_cbgs_pop) #pop data by year, cbgs levels

resident_pop_by_year <- 
  ca_cbgs_pop %>%
  filter(
    census_block_group 
    %in% substr(zip_blocks$GEOID10,1,12)) %>%
  group_by(YEAR = year) %>%
  summarize(
    resident_pop = sum(pop)
  )
saveRDS(resident_pop_by_year, "94108_resident_pop_by_year.rds")
```

## building energy consumption by job
```{r job-count, eval =F}
#get job counts data from LODES WAC in ZIP=94108 2013-2019
t <- read_csv("/Volumes/GoogleDrive/My Drive/Data Library/LODES/ca_wac_S000_JT00_2013.csv")

job_cnts <- 2013:2019 %>% 
  map_dfr(function(year){

    print(year)

    temp <- read_csv(paste0("/Volumes/GoogleDrive/My Drive/Data Library/LODES/ca_wac_S000_JT00_", year, ".csv")) %>%
      filter(w_geocode %in% zip_blocks$GEOID10) %>%
      mutate(
        year = year,
        job_cnts = C000
      ) %>%
      select(
        w_geocode,
        year,
        job_cnts
      )

    saveRDS(temp, paste0("94108_wac_", year, ".rds"))

    return(temp)

  }) %>%
  group_by(YEAR = year) %>%
  summarize(
    job_cnts=sum(job_cnts)
  )

saveRDS(job_cnts, "94108_job_total_by_year.rds")
```

```{r combine-job-res-energy-by-year, eval = F}
job_cnts <- read_rds("94108_job_total_by_year.rds")
res_pop <- read_rds("94108_resident_pop_by_year.rds")
names(job_cnts)

bldg_energy <- bldg_pge_by_year %>%
  ungroup() %>%
  left_join(job_cnts) %>%
  left_join(res_pop) %>%
  select(-ZIPCODE) %>%
  mutate(
    intensity = case_when(
      BUILDINGTYPE == "C" ~ TOTALKBTU/job_cnts,
      BUILDINGTYPE == "R" ~ TOTALKBTU/resident_pop
    )
  )

saveRDS(bldg_energy, "94108_bldg_energy_by_year.rds")
```

```{r building-energy-final}
bldg_energy <- read_rds("94108_bldg_energy_by_year.rds")
```

```{r G&E*C&R-line-plot}
bldg_energy %>% 
  ggplot(
    aes(
      x = YEAR,
      y = intensity
    )
  ) + 
  geom_line(
    aes(
      color = BUILDINGTYPE,
      linetype = ENERGYTYPE
    ),
    size = 1
  ) +
  labs(x = "Year", y = "kBTUs per resident or job", title = "ZIPCODE 94108 Annual Building Energy Use Intensity, 2013 to 2019", color ="Use Type", linetype = "Energy Type") +
  scale_linetype_manual(values = c("solid","dotted"), labels = c("Electricity","Gas"))
```

```{r HDD-CDD, eval = F}
cdd <- read_csv("CDD_SF_2013_2019.csv")
hdd <- read_csv("HDD_SF_2013_2019.csv")

HDD_CDD <- data.frame(
  YEAR = c(2013:2019),
  CDD = c(9,10,17,7,5,27,2),
  HDD = c(91,79,80,100,106,80,103)
)

bldg_energy_normalized <- bldg_energy %>%
  left_join(HDD_CDD) %>%
  mutate(
    intensity_normalized = case_when(
      ENERGYTYPE == "E" ~ intensity/CDD,
      ENERGYTYPE == "G" ~ intensity/HDD
    )
  )

saveRDS(bldg_energy_normalized, "94108_bldg_energy_normalized.rds")
```
